

## Environment

```{r, echo = TRUE}
### check if required libraries are installed 
packages <- c("dplyr", "xts")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())))  
}

### loading libraries
library(dplyr)
library(xts)

### read the dataset
setwd("C:/Users/user/Documents")
file <- read.table("learner_item_data.csv", header=TRUE, sep=",", row.names=NULL)
newdata <-arrange(file, uuid, created_at)


print(head(newdata))
```


## Functions

loading two major functions: 1. sorting dataset 2. session analysis


### Sorting function:

* input:  the column of time stamp

* output: user id / session starting time / time difference

### Session anlaysis function:

* input: xts format dataset

* output: analysis by date ( xts files in a list )


```{r, echo = TRUE}
check <- function(list){
	len <- nrow(list)
	matrix <- matrix(NA, ncol= 3, nrow = len)
	# user id, Date, length 

	start = list[1,2]
	for (num in 2:len){
		time = as.numeric(difftime(list[num,2], list[num-1,2], units = "mins"))

		# condition: allowed idle time < 10 mins, or users abort the session
		if (list[num-1,1] == list[num,1] && time < 10){matrix[num] = 0}
		else {
		matrix[num-1, 1] = list[num, 1] # register the id
		matrix[num-1, 2] = start
		matrix[num-1, 3] = as.numeric(difftime(list[num-1,2], start, units = "mins"))
		start = list[num,2] # beginning time of the next session 
		}
	}
	out <- na.omit(matrix)
	return (out)
}

session_analy <- function(days){
	len <- length(days)
	list = list() # output
	list_order = 1
	for (num in 1:len){
		table = table(days[[num]]$User_Id)
		multi_session <- length(table[table>1])

		# check if there's user using more than one session in a day (looping through)
		if (multi_session != 0 ){
				people<-as.xts(length(table[table>1]), order.by=index(head(days[[num]], n=1)))
				# number of people do several sessions a day

				temp <- as.xts(t(as.list(table[table>1])), order.by=index(head(days[[num]], n=1)) )
				# number of sessions by users who do several sessions a day

				final <- merge(people,temp)
				colnames(final) <- c("Active Users", dimnames(temp)[[2]] )
				list[[list_order]] <- final	 # append into a list
				list_order = list_order + 1
		}
	}
	return(list)
}

```



## Q1. Data Cleaning 

The idea is to identify the sessions of each user. I chose not to encode them in the dataset in a way, considering of using time-series based data format to split by dates. Here I create a customized check function. It takes one argument: sorted dataset by ID and time stamps.

1. register the first starting time as a variable called *start*

2. check two conditions (starting from 2nd)

    + if the current ID is the same as the previous ID? 
    
    + if the time difference is larger than 10 minutes? 
    
    + (not applied) check if a full session lengthis larger than 0.05 mins (3 sec)

  
3. if Yes, mark 0 and iterate

4. if no, we detect a session and write the information into a matrix

    + ID
    
    + start time of a session
  
    + total length of a session 
    
5. output the matrix


```{r, echo=TRUE}
cleaned <- check(newdata)
print(head(cleaned))
```


### Convert to times series dataset

```{r, echo=TRUE}
cleaned_ts <- xts(cleaned, as.Date(newdata[cleaned[,2],][,2]))
colnames(cleaned_ts) <- c("User_Id", "Starting_time", "Time_in_Mins")

print(head(cleaned_ts))
```


### 2.1 What is the distribution (histogram) of session lengths? 

```{r, echo=TRUE} 
X <- cleaned_ts[,3]
hist(X, prob=TRUE, col="grey")# prob=TRUE for probabilities not counts
lines(density(X), col="blue", lwd=2) # a density estimate with defaults
lines(density(X, adjust=2), lty="dotted", col="darkgreen", lwd=2) # "smoother" density
```

### 2.2 How many people do several sessions a day? How many sessions? 


#### Divided by dates

The reason of choosing *xts* is to enable the quick split at this step. Here, we order *split.xts* function to take the arguement: data and interval, and chunk the data into lists.

```{r, echo=TRUE} 
split_by_date <- split.xts(cleaned_ts, f="days") # split dataset by days 
print(head(split_by_date))
```

#### Final analysis 

We create a function (above) called *session_analy*,reorganizing the splited file by date. The function proceeds by the following steps:

1. Check whether there are IDs using mutitple sessions per day ($> 1$)

2. If there is (so *multi_session* is not zero), then 

    + 1st column: number of active users (using mutiple session per day)
  
    + From 2nd columns: IDs (attributes) and the times of sessions per day 
  
    + merge above into one *xts* file and store it into a list 

3. output the list


```{r, echo=TRUE} 
session_answer <- session_analy(split_by_date)
print(head(session_answer))
```
